Analyse der Survey-Daten von CHILDREN for a better World e.V.

### 2020-03-02 test.Rnw

% in Terminal enter
% Rscript -e "library(knitr); knit('paper.Rnw')"
% pdflatex paper
% biber paper
% pdflatex paper
% pdflatex paper

% IMPORTANT

% Install tex distribution
% https://www.latex-project.org/get/
% With Windows, install MikTeX

% in RStudio, install.packages("knitR")
% also, install all packages that appear in first code chunk in this file, i.e. in the one that looks like this
% <<LoadingPackages, include=FALSE>>=
% library(knitr)
% library(xtable)
% library(dplyr)
% library(stargazer)
% library(cowplot)
% @

% Go to Tools/Global Options/Sweave
% Select Weave Rnw files using: knitr
% Select Typeset LaTeX into PDF using: pdfLaTeX

% Use button "Compile PDF" to knit the Rnw file to a tex file and compile the tex file to a pdf file.

% Always install all packages that you're asked to after having pressed "Compile PDF".

% If there is no error while knitting from Rnw to tex, but an error while compiling from tex to pdf, have a look at the pdf file. It might still have worked.

% instead of Compile PDF, go into Terminal and enter 
% Rscript -e "library(knitr); knit('paper.Rnw')"
% latexmk -pdf paper

% turn page 90 degrees with lscape Gandrud (2016), p.185

% Create clickable hyperlinks with hyperref package, Gandrud (2016), p. 218

% Use natbib package for bibliography formatting, authoryear for Harvard style , Gandrud (2016), p. 227
% commands to change what information is included in the parentheses, Gandrud (2016), p. 228, table 11.2

% Using saveRDS and readRDS, following 
% https://csgillespie.github.io/efficientR/efficient-inputoutput.html#efficient-data-export-.rdata-or-.rds

% working with knitR and LaTeX
% https://kbroman.org/knitr_knutshell/pages/latex.html

% rotate tables in knitr
% https://stackoverflow.com/questions/21840878/rotate-a-table-from-r-markdown-in-pdf
% in particular stargazer tables
% https://stackoverflow.com/questions/31169845/rotate-stargazer-table-in-knitr/34292292#34292292

% Create list of packages used, Gandrud (2016), p 218/219

% models with robust standard errors texreg
% https://github.com/DeclareDesign/estimatr/issues/316

\documentclass[12pt, a4paper, titlepage]{article}
\usepackage{lscape}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
% bibstyle=apa, citestyle=authoryear
\usepackage[backend=biber, style=apa]{biblatex}
\DeclareLanguageMapping{english}{english-apa}
\usepackage{floatrow}
\usepackage{booktabs}
\usepackage[ngerman, english]{babel}
\usepackage{csquotes}
% \usepackage{bookmark}
\usepackage{rotating}
\usepackage{dcolumn}
% \usepackage[authoryear]{natbib}
% \bibliographystyle{apalike}
\usepackage{mathtools}
% \usepackage[onehalfspacing]{setspace}
\usepackage[left = 2.5cm, top = 2.5cm, right = 2.5cm, bottom = 2.0cm]{geometry}
\usepackage[hidelinks]{hyperref}

% 1.5 spacing in Word is about this in LaTeX
% https://tex.stackexchange.com/questions/65849/confusion-onehalfspacing-vs-spacing-vs-word-vs-the-world

\linespread{1.25}

\addbibresource{references.bib}

@@

<<CentAlignFig, include=FALSE>>=
# Center align all knitr figures
# Gandrud (2016), p. 53-55
opts_chunk$set(fig.align = "center",
                out.width = "0.8\\textwidth")
@

% multiple authors
% https://tex.stackexchange.com/questions/4805/whats-the-correct-use-of-author-when-multiple-authors

\title{Analysis of Survey Data Collected by CHILDREN for a better World e.V.}
\author{
Laura Huber\\
%\texttt{11709507}
\and
Laura Jepsen\\
%\texttt{11567799}
\and
Jonathan Kirschner\\
%\texttt{11724530}
\and
Rafael Schütz\\
%\texttt{11574909} 
\and 
Yannick Zurl\\
%\texttt{11701793}
\and
Studentisches Praxisprojekt zur Empirischen Wirtschaftsforschung PaRE3To\\
\and
Ludwig-Maximilians-Universität Munich
}

% Winter Term 2019/20\\

\date{27th February 2020}

@@

% \begin{titlepage}
% \maketitle
% \end{titlepage}

\tableofcontents
\listoftables

% note below table
% https://stackoverflow.com/questions/44549464/table-spacing-issue-converting-to-pdf-via-latex-with-pandoc/44797387#44797387
\listoffigures

\begin{abstract} 
This study describes the analysis of the data from the non-profit organisation 'CHILDREN for a better world e.V.'. We obtain null-results associated to the subsidy CHILDREN provides for its two flagship programs, Meals and Trips, and selected outcome variables. We find a strong but only partially robust connection between health relevant outcomes and the index of healthy diet criteria fulfilled in organization's menus. Using a differences-in-differences strategy, we cannot show a significant effect of participating in the Trips program on the everyday expertise and selfworth of the beneficiaries.
\end{abstract}

\section{Introduction}
CHILDREN for a better World e.V. is a non-profit organization supporting social institutions for children and adolescents across Germany, such as youth centers. Their goal is to foster equality of opportunities for the youth in Germany. CHILDREN supports poor and disadvantaged families and young people in socially problematic areas. Furthermore, they fight child poverty and encourage the youth to engage in society. We call the institutions funded by CHILDREN organizations in the following. Organizations have to apply for funding from CHILDREN each year again. In the application process, CHILDREN requires the applicants to answer a survey which contains questions about outcomes related to CHILDREN’s programs as well as general information about the organization such as the yearly budget.
CHILDREN provided us with data from two of its flagship programs: Mittagstisch (we refer to this as Meals program) and Entdeckerfonds (Trips program). The organizations use money from the Meals program to finance meals, from breakfast to dinner, that they sell at concessionary prices to the children and adolescents that visit them. In the following, we call these children and youth who ultimately profit from CHILDREN's grants beneficiaries. The organizations also use money from the Trips program to make trips to nearby places usually unknown to the beneficiaries.

Previous research shows that childhood and youth are central in development of character traits and future opportunities. Factors like education, stable family conditions, social contacts and mobility shape children and adolescents for their further life. Children from parents with lower income and lower educational attainment have typically a lower weight at birth and are born in a later week of gestation \parencite{Deckers.27.04.2015}, which are indicators for bad initial conditions, as shown by \textcite{Case.2002}. Children that grow up in families with a low socio-economic status are often less mobile and more likely to stay amongst themselves. This could have a negative influence on their character development. For example, \textcite{Chetty.2014} investigate for the US, that areas with more mobility are highly correlated with better primary schools, greater social capital and more stable family conditions. Furthermore, the result of \textcite{Kosse.2020} provides causal evidence on the effect of social environment on prosocial attitudes. Being supported by a mentor could result in a significant and persistent increase in the prosociality of elementary school children, with regard to prosocial role models and intense social interactions. \textcite{Heckman.2010} investigated a preschool education program in the US. He shows that an investment in the improvement of the childhood conditions could have a high rate of return, even by controlling for possible distortions.

The aim of our project paper is to test whether CHILDREN’s programs positively influence the beneficiaries. We look at the associations between the subsidy CHILDREN provides and the share of beneficiaries with broadened everyday expertise and improved selfworth. Furthermore, we analyze the association between the measure of the healthiness of the meals the organizations offer and health-relevant outcomes of beneficiaries. The aim of our applied differences-in-differences estimation is to test whether the activities offered by the Trips program have a positive effect on the participating children, measured through a change in selfworth and everyday expertise.
We find no profound connection between the subsidy CHILDREN provides and the effectiveness of the organizations regarding the share of beneficiaries with broadened everyday expertise and improved selfworth. On the other hand, focusing on health relevant variables, we obtain a positive association between the index of healthy diet criteria fulfilled by organizations and the share of beneficiaries who are less frequently ill. By using the differences-in-differences (DID) strategy, the estimates show that the participation in the Trips program has a positive, but insignificant effect on the everyday expertise of beneficiaries. Strikingly, we observe a negative association between participating in the Trips program and selfworth. However, the results of the DID estimation should not be overstated because of the small sample size.
The remainder of this project paper is structured as follows: Section \ref{data} describes the variable used for the empirical analysis. Then we present summary statistics. We explain the dimensionality reducing method factor analysis. The proceeding section reports different regression results. After that, we implement the differences-in-differences strategy to test for the effects of the Trips program. Finally, we conclude and provide suggestions for CHILDREN.

\section{Data}
\label{data}
To measure the effect of CHILDREN's programs on the supported organizations, we use data from 2011 to 2018. In each year CHILDREN sends a survey to the organizations with several questions about the previous year. The number of organizations varies across years and increases over time: From 52 organizations in 2011 to 68 organizations in 2018. In some organizations one specific employee answers the survey questions, while in other institutions the questionnaire is answered in a team. As the children and adolescents are not questioned directly, all responses are documented as perceptions of the employees. The constellation of the variables in the data set also varies across years. For instance, CHILDREN asks for the average amount of kids that increased their confidence or improved their dietary knowledge. These questions of the survey are based on a scale of five levels from no to all kids. We perform several steps to receive a full dataset, which we use in our empirical analysis. Originally, the data was divided into seperate data sets for the respective years. We only use the surveys until 2018, because in 2019 some organization-ID's occurred multiple times and the dataset for 2019 was incomplete. To ensure comparability across years, we named all variables the same. We merged the different datasets to a common dataset, including all years and variables CHILDREN collected. For an efficient and clear data structure, we created a function that automatically changed the data type of all variables from "character" to "ordinal" and added several versions for each initially metric encoded variable afterwards. The three variants are ordinal, standardized and weighted.
Furthermore, we created a set of new variables. For instance, we used the information CHILDREN gave us to assign the German states to the corresponding organization-ID. 
The final dataset that we used for the empirical analysis is structured as follows: Each row represents a specific organization observed in a specific year. The questions are divided in two categories. The first part of the data set includes variables regarding the Meals program. The second part contains variables affecting the Trips program. The final data set comprises the years 2011 to 2018.
Many organizations do not answer all questions CHILDREN poses. We create a seperate data set, in which we impute missing values with an organization-specific linear trend. CHILDREN supports some very large organizations that give out hundreds of meals a day or conduct dozens of trips per year. We fit our models with outliers excluded to see if they are driving results. We define an outlier as a value that is at least 1.5 interquartile ranges below the 25th percentile or equally far above the 75th percentile. Once we exclude outliers in terms of numbers of meals and once in terms of number of trips.

\section{Summary Statistics}
\subsection{Fundamental Dynamics} 
In this section, we give an overview of the dynamics of CHILDREN's two flagship programs. We focus on the number of estimated ultimate beneficiaries in both programs, median total subsidy, median subsidy per institution, and median subsidy per beneficiary. We also look at selected outcomes, i.e. those related to health as well as self-worth and day-to-day skills. We have converted all nominal monetary variables into 2015 euros, using price indices from the Federal Statistical Office of Germany (Statistisches Bundesamt). We deflate (requested) grants as well as organizations' total expenses for the Meals program  with the price index related to food and non-alcoholic beverages (in German: Nahrungsmittel und alkoholfreie Getränke) and (requested) grants towards the Trips program with the price index for leisure, entertainment, and culture (in German: Freizeit, Unterhaltung und Kultur). These are only available after logging in on DESTATIS. The organizations also gave information about their total yearly budget. We inflate this with the general price index.

<<FundamentalDynamics, echo=FALSE, results='asis', message=FALSE, warning=FALSE>>=

@@

fundamental_dynamics = readRDS("./ANALYSIS/Tables/fundamental_dynamics.Rds")

colnames(fundamental_dynamics) <- c("Year", "Beneficiaries, Meals", "Beneficiaries, Trips", "Organizations, Meals", "Organizations, Trips")

xtable(fundamental_dynamics,
                                  # summary = FALSE,
                                digits = c(0, 0, 1, 1, 0, 0),
           caption = "Summary Statistics",
           align = "lccccc", 
          label = "fundamentalDynamics"
          )

@
% look up official names of price indices in English

Table \ref{fundamentalDynamics} shows that, at the beginning of the time series in 2011, in the Lunch program they supported in 52 organizations. In 2018, this number has increased to 68. In 2011, CHILDREN financed meals for 3748 beneficiaries, and for 5103 in 2018. In the Trips program, which launched in 2012, the number of supported organizations amounted to 44 in 2012 and grew to 49 until 2018. In 2012, their grant allowed 2803 beneficiaries to go on trips, and 6911 in 2018.

\subsection{Trends of Grants}

\begin{figure}
  \caption{Yearly Dynamics of Total Grants in Meals and Trips Program}
  \label{totalGrantsDyn}

<<GrantTrend, echo=FALSE, message=FALSE>>=
totalGrantsDyn = readRDS("./ANALYSIS/GRAPHS/PAPER/summaryStatistics_Subsidy.Rds")
print(totalGrantsDyn)
@

\floatfoot{This graph shows the development of grants in the Meals compared to the Trips program. We distinguish between the sum of grants in one year, the median grant and the median grant per beneficiary. From left to right: Meals, Trips. From top to bottom: sum, median, median per beneficiary. We have deflated the values to 2015 euros using the price index related to food and non-alcoholic beverages(in German: Nahrungsmittel und alkoholfreie Getränke) for the Meals program and the price index related to Leisure, Entertainment and Culture (in German: Freizeit, Unterhaltung, Kultur) provided by the Federal Statistical Office of Germany (Statistisches Bundesamt).}
\end{figure}

For the dynamics of figure \ref{totalGrantsDyn} we visualze the dynamics of CHILDRENs' grants by distinguishing between the sum of grants in one year, the median and the median grant per beneficiary. We also compare the grants of the Lunch and the Trips program.

Figure \ref{totalGrantsDyn} shows no exact trend for the total real grant for the Lunch program. Between 2013 and 2015 the grant increased from about 600000 EUR to 680000 EUR, but falls behind in the following year. Since 2017 an increase is again visible. In comparison to this, there is a clearer negative trend in median meals subsidy, falling from about 8000 EUR in 2012 to about 6500 EUR in 2018. In accordance, the median meals grant per beneficiary shrinks from about 250 EUR to about 200 EUR.

In the total trips grant, a slightly negative trend is visible, after the subsidy increased to approximately 200000 EUR in 2014, but decreased to about 130000 EUR in 2016, and is connstand eversince. The trips median grant therefore also decreased from 3000 EUR to 2000 EUR over the time period, as well as the median grant per beneficiary decreased from about 60 EUR to about 30 EUR.

These results visualize, for the Lunch as well as for the Trips program, together with the fundamental dynamics of table \ref{fundamentalDynamics}, that CHILDREN was able to increase the number of organizations they support, but had to distribute the available subsidy among them.

\subsection{Dynamics of Health-Related Variables}

In its yearly surveys, CHILDREN asks about three variables closley related to a healthy diet. These are the shares of beneficiaries who are healthier, have a growing appreciation for a healthy diet, or have increased their knowledge about what constitutes a healthy diet. Figure \ref{HealthTimeplots} displays the share of organizations in each category of the health outcome from 2011 to 2018. The possible values are: all, most, some, few, and none. For this figure, we use the original ordinal varibles which result from the survey structure.

\begin{figure}
  \caption{Health Outcome Over Time}
  \label{HealthTimeplots}
<<HealthTimePlots, echo=FALSE, message=FALSE>>=
Health_timeplots <- readRDS("./ANALYSIS/GRAPHS/PAPER/Health_Year.Rds")
print(Health_timeplots)
@
\floatfoot{In its yearly surveys CHILDREN asks about three variables closley related to a healthy diet. These are the shares of beneficiaries who are less frequently ill (lessIll\_ordered), with expanded dietary knowledge (dietaryKnowledge\_ordered), or with increased appreciation for a healthy diet (appreciateHealthy\_ordered).
The x-axis plots the year. The y-axis displays the share of organizations in each category of the health outcome. The possible values are: all, most, some, few and none.}
\end{figure}

In Figure \ref{HealthTimeplots}, for the variable 'lessIll\_ordered', there is much non available data (lessIll\_ordered refers to the share of beneficiaries who are less frequently ill). In all years, most organizations say that most beneficiaries are more healthy. The least stated category is, that all beneficiaries are less ill. That few are less ill, appears sometimes and the leftover category, none, which is only coded for the variable 'lessIll\_ordered', appears as good as never.

In the second plot of figure \ref{HealthTimeplots} (variable 'dietaryKnowledge\_ordered', refers to the share of beneficiaries with expanded dietary knowledge) most organizations state, that predominantely most or all beneficiaries increased their knowledge about what constitutes a healthy diet. The least stated category is that some beneficiaries increased their dietary knowledge. The leftover category, few, does not appear.

The bottom plot of figure \ref{HealthTimeplots}, which visualizes the variable 'appreciateHealthy\_ordered' (refers to the share of beneficiaries with increased appreciation for a healthy diet). Again most organizations state, that most beneficiaries have a growing appreciation for a healthier diet. The second most stated answer is that all beneficiaries have a growing appreciation for a healthier diet. The least stated category by the organizations is, as well as in the second plot, that some have a growing appreciation for a healthier diet. The leftover category, few, also does not appear.

\subsection{Selfworth and Everyday Expertise}

In its yearly surveys, CHILDREN has always asked about the following two variables. These are the shares of beneficiaries with improved self-worth and with broadened everyday expertise. Figure 4 displays the share of organizations in each category of the health outcome from 2011 to 2018. The possible values are: all, most, some, few, and none. As well as in the last figure (figure 3), we use the original ordinal varibles which result from the survey structure.

\begin{figure}
  \caption{Selfworth and Everyday Expertise Over Time}
  \label{EqualityTimeplots}
<<EqualityTimePlots, echo=FALSE, message=FALSE>>=
Equality_timeplots <- readRDS("./ANALYSIS/GRAPHS/PAPER/Equality_Year.Rds")
print(Equality_timeplots)
@
\floatfoot{In its yearly surveys CHILDREN has always asked about following two variables. These are the shares of beneficiaries with improved self-worth (selfworth\_ordered) and with a broadened understanding for everyday expertise (dayToDaySkills\_ordered).
The x-axis plots the year. The y-axis displays the share of organizations in each category of the health outcome. The possible values are: all, most, some, few, and none.}
\end{figure}

Figure \ref{EqualityTimeplots} shows in its upper plot the development of the answered categories for the variable selfworth\_ordered (refers to the share of beneficiaries with improved self-worth). Mostly, the answer is that most beneficiaries in most organizations have more selfworth. The second most stated category is that all beneficiaries gained in selfworth and the least stated is that some have more selfworth. The leftover category, few, does not appear.

In the bottom plot of figure \ref{EqualityTimeplots} the development of the share of beneficiaries with a broadened understanding for everyday expertise over time is visible. The results are about similar to the previous results of selfworth: Most beneficiaries in most organizations have a growing understanding of everyday expertise.

\section{Feature Selection}

CHILDREN survey many variables, some of which seem closely related to each other. We use exploratory factor analysis for dimensionality reduction and employ extracted factors as controls in the models we estimate. In the appendix, we use the partition method suggested by \textcite{Millstein.2020} to establish a one-to-one correspondence between apparent and underlying variables.

\subsection{Exploratory Factor Analysis}

The fundamental theorem of factor analysis posits that each observable variable can be expressed as a linear combination of a smaller number of underlying factors as well as an error term. The coefficients of the factors are called loadings. \parencite[p.310]{Moosbrugger.2008}

For factor analysis, we work with the imputed data set. We conduct a separate factor analysis for each model that we want to fit. Each time, we include all variables that were recorded in all years the Meals and the Trips program have been run respectively, but exclude those that we want to use as response and predictor in the corresponding model.

First, we determine the optimum number of factors via the stochastic approach of parallel analysis, following the advice of \textcite[p.313]{Moosbrugger.2008}.
We estimate factor scores and loadings with the Bartlett method, which is similar to a Maximum Likekelihood estimation, and presented as the most convential one in \textcite[p.291]{Eid.2014}.
As the fundamental theorem of factor analysis has no unique solution, we have to decide on how to rotate the factors. As we want them to be orthogonal to each other, we use the rotation technique Varimax, which is recommended for this purpose in \textcite[p.317]{Price.2017}.

% \subsection{Double Selection}

\section{Associations Between CHILDREN's Grants and Selected Outcomes}

\subsection{Empirical Approach}

In this section, we use a simple linear model, as described in equation \ref{SimpleLinearModel}. We estimate each model with the original data set, the one with imputed variables, and the one with excluded outliers. When appropriate, we include the relevant extracted factors as controls.

\begin{equation}
\label{SimpleLinearModel}
  y_{it} = \beta_0 + \beta_1 x_{it} + \epsilon_{it}
\end{equation}

 We look at the association between:

\begin{itemize}
  \item{the subsidy (in 2015 EUR) an organization receives through CHILDREN's Meals program and the number of meals it dispenses}
  \item{the subsidy (in 2015 EUR) an organization receives through CHILDREN's Trips program and the number of trips it conducts}
  \item{the subsidy per beneficiary (in 2015 EUR) an organization receives through CHILDREN's Meals program and the standardized share of beneficiaries with improved self-worth}
  \item{the subsidy per beneficiary (in 2015 EUR) an organization receives through CHILDREN's Meals program and the standardized share of beneficiaries with broadened everyday expertise}
   \item{a standardized measure of the healthiness of the meals an organization dispenses and various standardized health-related outcomes of beneficiaries} 
\end{itemize}

We discuss all these models in turn.

\subsection{Direct Effects of CHILDREN's Grants}

If CHILDREN's grants are to have an effect on beneficiaries, they should first influence the output of organizations in terms of meals dispensed and trips conducted.

<<GrantsRegressionsLunch, echo=FALSE, results='asis', message=FALSE>>=

lm_ms = readRDS("./ANALYSIS/Tables/mealsNo_sub.lm.Rds")
lm_ms_c = readRDS("./ANALYSIS/Tables/mealsNo_sub_control.lm.Rds") #control
lm_ms_o = readRDS("./ANALYSIS/Tables/mealsNo_sub_ex.lm.Rds")#outlier meals
lm_ms_o_c = readRDS("./ANALYSIS/Tables/mealsNo_sub_ex_control.lm.Rds")#outliercontr
lm_ms_i = readRDS("./ANALYSIS/Tables/mealsNo_sub_IM.lm.Rds")#impute meals

texreg(list(lm_ms, lm_ms_c, lm_ms_o, lm_ms_o_c, lm_ms_i), caption = "Association Between Number of Meals and Real Subsidy", custom.model.names = c("(1)", "(2)", "(3)", "(4)", "(5)"), scalebox = 0.8, custom.note = "\\parbox{\\linewidth}
{\\vspace{2pt} Dependent variable: number of meals \\\\ realSubsidy: subsidy for Meals program in 2015 EUR \\\\ eatersPerMeal: number of beneficiaries of Lunch program \\\\ Model (1): original data set, simple linear model, estimated with OLS \\\\ Model (2): original data set, linear model with controls, estimated with OLS \\\\ Model (3): data set without outliers, simple linear model, esmitaed with OLS \\\\ Model (4): data set without outliers, linear model with controls, estimated with OLS \\\\ Model (5): imputed data set, simple linear model, estimated with OLS \\\\ All regressions are estimated with robust standard errors %stars.}", caption.above = TRUE, label = "GrantsRegressionsLunch", booktabs = TRUE, use.packages=FALSE)

@

As table \ref{GrantsRegressionsLunch} shows, this is emphatically true. Whether we look at the original data set, the one without outliers or the one with imputed values, a strong association becomes evident. 
The estimated coefficients are very similar when we use the original dataset and the one with imputed values. In the case of the original data set, increasing the subsidy of an organization by one EUR is associated with 2.6 additional meals dispensed. This estimate is highly statistically significant. When we exclude outliers, i.e. those organizations that give out very many menus or very few menus,
the estimated coefficient decreases by about one order of magnitude. Still, spending not much more than three EUR more goes hand in hand with one extra meal. The estimate is also highly statistically significant.

In the remit of the Trips program, the picture looks much different. In the case of the original data set and the data set with imputed variables, an increase in the trips subsidy by 5000 EUR is needed for an extra trip. The coefficent of 0.0002 is statistically significant at the 5\% level. As soon as outliers are excluded, the coefficent changes to 0.0003, which is highly statistically significant. This means that, according to this model, at least 3000 EUR are needed for an additional trip.
In sum, there seems to be no clear connection between how much money CHILDREN gives to an organization and the number of trips it organizes.

<<GrantsRegressionsTrips, echo=FALSE, results='asis', message=FALSE>>=
lm_ts = readRDS("./ANALYSIS/Tables/tripsNo_sub.lm.Rds")
lm_ts_c = readRDS("./ANALYSIS/Tables/tripsNo_sub_control.lm.Rds")
lm_ts_o = readRDS("./ANALYSIS/Tables/tripsNo_sub_ex.lm.Rds") #outlier trips
lm_ts_o_c = readRDS("./ANALYSIS/Tables/tripsNo_sub_ex_control.lm.Rds")
lm_ts_i = readRDS("./ANALYSIS/Tables/tripsNo_sub_IM.lm.Rds") #imputetrips

texreg(list(lm_ts, lm_ts_c, lm_ts_o, lm_ts_o_c, lm_ts_i), caption = "Association Between Number of Trips and Real Subsidy", custom.model.names = c("(1)", "(2)", "(3)", "(4)", "(5)"), scalebox = 0.8, custom.note = "\\parbox{\\linewidth}
{\\vspace{2pt} Dependent variable: number of trips \\\\ realTripsSubsidy: subsidy for Trips program in 2015 EUR \\\\ tripsKidsNo: number of beneficiaries of Trips program \\\\ Model (1): original data set, simple linear model, estimated with OLS \\\\ Model (2): original data set, linear model with controls, estimated with OLS \\\\ Model (3): data set without outliers, simple linear model, esmitaed with OLS \\\\ Model (4): data set without outliers, linear model with controls, estimated with OLS \\\\ Model (5): imputed data set, simple linear model, estimated with OLS \\\\ All regressions are estimated with robust standard errors %stars.}", caption.above = TRUE, label = 'GrantsRegressionsTrips', digits = 4, booktabs = TRUE, use.packages=FALSE)
@

\subsection{Variables of Interest: Selfworth and Everyday Expertise}

In their surveys, CHILDREN asks about two variables in both programs. These are the share of beneficiaries with improved self-worth and share of beneficiaries with broadened everyday expertise. This feature could potentially allow us to compare the two programs regarding their realtive effectiveness. Like always, we standardize the two outcome variables. As before, we can only make associational claims. We invariably find no clear-cut relationship between the subsidy per beneficiary and the two outcomes. In fact, no coefficent is statistically significantly different from zero at any of the usual levels. 
The four variables are recorded for each year in which the program was operative.
Tables \ref{SelfworthRegressions} and \ref{DayToDaySkillsRegressions} present the null results.

<<SelfworthRegressions, echo=FALSE, results='asis', message=FALSE>>=

ss = readRDS("./ANALYSIS/Tables/selfworth_scaled.Rds") 
tss = readRDS("./ANALYSIS/Tables/selfworthTrips_scaled.Rds")#sw
ssI = readRDS("./ANALYSIS/Tables/selfworth_IM_scaled.lm.Rds") #standardized
tssI = readRDS("./ANALYSIS/Tables/selfworthTrips_IM_scaled.lm.Rds") 
sfactor = readRDS("./ANALYSIS/Tables/factorselfworthmeals.Rds")

texreg(list(ss, tss, ssI, tssI, sfactor), caption = "Association Between Selfworth and Subsidy per Beneficiary", custom.model.names = c("(1)", "(2)", "(3)", "(4)", "(5)" ), scalebox = 0.8, custom.note = "\\parbox{\\linewidth}
{\\vspace{2pt} realSubsidyPerBeneficiary: subsidy per beneficiary of Meals program in 2015 EUR \\\\ realTripsSubsidyPerBeneficiary: subsidy per beneficiary of Trips program in 2015 EUR \\\\Model (1): dependent variable: share of beneficiaries with improved self-worth in the Lunch program, original data set, simple linear model, estimated with OLS \\\\ Model (2): dependent variable: share of beneficiaries with improved self-worth in the Trips program, original data set, simple linear model, estimated with OLS \\\\ Model (3): dependent variable: share of beneficiaries with improved self-worth in the Lunch program, imputed data set, simple linear model, estimated with OLS \\\\ Model (4): dependent variable: share of beneficiaries with improved self-worth in the Trips program, imputed data set, simple linear model, estimated with OLS \\\\ Model (5): dependent variable: share of beneficiaries with improved self-worth in the Lunch program, imputed data set, linear model with extracted factor scores as controls, estimated with OLS \\\\ All regressions are estimated with robust standard errors %stars.}", caption.above = TRUE, label = 'SelfworthRegressions', bookmarks = TRUE)
@

<<DayToDaySkillsRegressions, echo=FALSE, results='asis', message=FALSE>>=
ds = readRDS("./ANALYSIS/Tables/dayToDaySkills_scaled.lm.Rds") #standardizedweigh
tds = readRDS("./ANALYSIS/Tables/dayToDaySkillsTrips_scaled.lm.Rds")
dsI = readRDS("./ANALYSIS/Tables/dayToDaySkills_IM_scaled.lm.Rds") 
tdsI = readRDS("./ANALYSIS/Tables/dayToDaySkillsTrips_IM_scaled.lm.Rds")
dfactorm = readRDS("./ANALYSIS/Tables/factordaytodayskillsmeals.Rds")
dfactort = readRDS("./ANALYSIS/Tables/factordaytodayskillstrips.Rds")

texreg(list(ds, tds, dsI, tdsI, dfactorm, dfactort), caption = "Association between everyday expertise and subsidy per beneficiary", custom.model.names = c("(1)", "(2)", "(3)", "(4)", "(5)", "(6)"), scalebox = 0.8, custom.note = "\\parbox{\\linewidth}
{\\vspace{2pt} realSubsidyPerBeneficiary: subsidy per beneficiary of Meals program in 2015 EUR \\\\ realTripsSubsidyPerBeneficiary: subsidy per beneficiary of Trips program in 2015 EUR \\\\ Model (1): dependent variable: share of beneficiaries with broadened everyday expertise in the Lunch program, original data set, simple linear model, estimated with OLS \\\\ Model (2): dependent variable: share of beneficiaries with broadened everyday expertise in the Trips program, original data set, simple linear model, estimated with OLS \\\\ Model (3): dependent variable: share of beneficiaries with broadened everyday expertise in the Lunch program, imputed data set, simple linear model, estimated with OLS \\\\ Model (4): dependent variable: share of beneficiaries with broadened everyday expertise in the Trips program, imputed data set, simple linear model, estimated with OLS\\\\ Model (5): dependent variable: share of beneficiaries with broadened everyday expertise in the Lunch program, imputed data set, linear model with extracted factor scores as controls, estimated with OLS \\\\ Model (6): dependent variable: share of beneficiaries with broadened everyday expertise in the Trips program, imputed data set, linear model with extracted factor scores as controls, estimated with OLS \\\\ All regressions are estimated with robust standard errors %stars.}", caption.above = TRUE, label = 'DayToDaySkillsRegressions', bookmarks = TRUE)

@

\subsection{Variables of Interest: Health Variables}

Now, we turn to a predictor that CHILDREN does not directly influence until now, but that it easily could. In 2014, 2016, 2017 and 2018 the organizations had to send CHILDREN a sample of their menus. An ecotrophologist collaborating with CHILDREN assessed those menus with regards to how healthy they were. This assessment was based on criteria of the German Nutrition Society (Deutsche Gesellschaft für Ernährung).

In its yearly surveys, CHILDREN asks about three variables closely related to a healthy diet. These are the shares of beneficiaries who are less frequently ill, with expanded dietary knowledge and with increased appreciation for a healthy diet.

Figure \ref{HealthPlots} gives an overview of the relationship between the healthy food criterion (DGECriteriaNo, index of healthy diet criteria fulfilled in organization's menu) and each of the three health-related variables. For these results, we use their ordinal values. The x-axis displays the index for a healthy diet. The y-axis displays the share of organizations in each category of the health outcome. The possible values are: all, most, some, few and none. For example, if an organization says that most beneficiaries are healthier, then this would be coded as most. In the last reporting period, there are 13 healthy diet criteria. In previous years there were up to 15. The ecotrophologist assessed each criterion individually and decided wether it was fulfilled or not. The healthy food index is calculated as the number of fulfilled criteria. Examples for criteria are daily carbohydrates or daily vegetables.

All plots of Figure \ref{HealthPlots} show that most organizations suffice about 5 to 10 healthy food criteria. The largest part of criterion values are associated with most beneficiaries being less ill, having a growing appreciation for a healthy diet or having increased their knowledge about what constitutes a healthy diet. It secondly becomes visible that the healthy criterion only exchanges with a growing appreciation for a healthy diet or an increasing dietary knowlege for all beneficiaries (which is the second most stated criterion in the two bottom plots), but not with all beneficaries being more healthy. However, for the variable less ill, the least data is availabe. Overall, the criteria, even a few, seem to have a positive relationship with beneficiaries health relevant variables. This connection is empirically examined in regressions \ref{HealthRegressions-LessIll}, \ref{HealthRegressions-DietaryKnowledge} and \ref{HealthRegressions-AppreciateHealthy}.

\begin{figure}
  \caption{Health Outcomes versus Healthy Meals}
  \label{HealthPlots}
<<HealthRegressionsPercentageplots, echo=FALSE, message=FALSE>>=
Health_percentageplots <- readRDS("./ANALYSIS/GRAPHS/PAPER/Health_percentageplots.Rds")
print(Health_percentageplots)
@
\floatfoot{DGECriteriaNo is an index of healthy diet criteria fulfilled in organization's menu. It is based on criteria of the German Nutrition Society (Deutsche Gesellschaft für Ernährung). According to information from CHILDREN, they ask the organizations to send them a sample of their menus. An ecotrophologist collaborating with CHILDREN assessed the menus. In its yearly surveys CHILDREN asks about three variables closley related to a healthy diet. These are the shares of beneficiaries who are less frequently ill (lessIll\_ordered), with expanded dietary knowledge (dietaryKnowledge\_ordered), or with increased appreciation for a healthy diet (appreciateHealthy\_ordered).
The x-axis plots the index for a healthy diet. The y-axis displays the share of organizations in each category of the health outcome. The possible values are: all, most, some, few, and none}
\end{figure}

We use standardized versions of the health related variables as outcomes in simple linear models with a standardized version of the healthy food criterion as a predictor in each.
By estimating these models with ordinary least squares (OLS), we ascribe the same weight to an organization where a thousand beneficiaries regularly eat as to one where only ten people do so. To control for this difference in size, we additionally fit the models with weighted least squares (WLS), using the number of beneficiaries as weights.

In Table \ref{HealthRegressions-LessIll}, after weighting the regression for the original data set, the coefficent rises slightly but is becoming less statistically significant. An increase of the healthy food criterion by one standard deviation is associated with an increase in beneficiaris being less frequently ill by 0.35 standard deviations. Regressing with the imputed dataset, results decrease by about 0.1. Using weighted least squares additionaly results in a slighty decreased coefficent, which is not significant anymore. Regressing with the imputed data set with extracted factor scores as controls, we still obtain a slightly postive significant coefficent: An increase of the healthy food criterion by one standard deviation is associated with an increase in beneficiaris being less frequently ill by 0.18 standard deviations. Overall, there is a positive association between the two variables.

An expanded model with the dependent variable less Ill is presented in Appendix 1 in table \ref{expandLessIll}.

>>>>

% Rscript -e "library(knitr); knit('test.Rnw')"
% pdflatex test
% biber test
% pdflatex test
% pdflatex test

\documentclass[12pt, a4paper, titlepage]{article}

@@

@@

<<Partition, echo=FALSE, results='asis', message=TRUE, warning=TRUE>>=

@@

partition_meals = readRDS("./ANALYSIS/Tables/PartitionMeals.4.Rds")
partition_trips = readRDS("./ANALYSIS/Tables/PartitionTrips.4.Rds")
